### 2024-03-16 paper/paper.qmd
finished final touches, waiting for posit cloud to become available for rendering

---
title: "Minimal Positive Influence of Helpline Efforts on Reduction of Homeless Death in Toronto"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/help-line-homeless-death-analysis.git"
date: today
date-format: long
abstract: "This paper analyzes the influence of Central Intake Line staff efforts on the reduction of homeless death count in Toronto. Data on coded number of calls from Central Intake Call Wrap-Up Codes dataset and data on monthly homeless death count from Deaths of People Experiencing Homelessness dataset are used in generalized linear regression models to estimate the influence. We find that the staffs’ net efforts create a statistically insignificant influence on the reduction of homeless death, with specific types of efforts appearing counterproductive. This implies that while a positive influence does exist, from a statistical point of view, the impact of the Central Intake Line staff is not significantly different from non-existent."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(readr)
library(rstanarm)
library(knitr)
library(modelsummary)
library(ggplot2)
```

# Introduction

The City of Toronto has been consistently providing funding and staff in
an effort to reduce the active homeless population within its governing
district. Amongst numerous city efforts, the 24/7 helpline — Toronto
Central Intake Line — was created to offer referrals to emergency
shelters, sleeping accommodations, and provide general information about
homelessness and prevention. [@Central_Intake_Call]. Starting in 2020,
detailed data about the Central Intake Line became available on the
OpenDataToronto portal [@OpenDataToronto].

The estimand of the paper focuses on examining the influence of helpline
efforts on the reduction of homeless deaths in Toronto. Using Central
Intake Call Wrap-up Codes data [@Central_Intake_Call] from
OpenDataToronto [@OpenDataToronto], three different aspects of the
helpline efforts were isolated: total calls coded, referral to shelter,
and homelessness prevention information provided. These three key
aspects were measured against the monthly cumulative death count from
Deaths of People Experiencing Homelessness Data [@Homeless_Death_Count]
documented by the Toronto Public Health [@TPH]. The analyses yielded an
insignificant negative correlation between the net efforts of the
helpline staff. The variable referral to shelter yields a positive
correlation with homeless death count, whereas the effort to provide
homelessness information yields a negative correlation. This implies
that the net influence of helpline efforts is statistically negligible,
while individual efforts can be productive or counterproductive. This
examination in the net efforts and individual efforts can provide the
helpline with a suggestion of which types of services or calls would be
more economically efficient to allocate their labor resources on.

Analyses and findings in this paper are structured into several
sections: @sec-data -- Data, @sec-model – Model, @sec-results --
Results, and @sec-discussion -- Discussion. The Data section examines
all datasets and variables kept for analysis, followed by an explanation
of their data cleaning processes. The Model section defines linear
models used for further analysis, explain its components, and presents
model justifications. The Result section focuses on visualizing and
presenting the model results through data presented in Data section. The
Discussion section further evaluate the interpretations behind the model
results presented in the previous section, and touches on any weaknesses
and next steps.

# Data {#sec-data}

All data used in this paper are obtained through OpenDataToronto Portal
[@OpenDataToronto]. Two different datasets: Central Intake Call Wrap-Up
Codes Data [@Central_Intake_Call] and Deaths of People Experiencing
Homelessness[@Homeless_Death_Count], are retrieved to analyze the effect
of Toronto’s Central Call Line (denoted as *Helpline* for the remainder
of this paper) efforts to the death counts of homeless individuals in
Toronto. Data is cleaned and analyzed using the open source statistical
programming language R [@citeR] and supporting packages tidyverse
[@tidyverse], janitor [@janitor], rstanarm [@rstanarm], arrow [@arrow],
ggplot2 [@ggplot2], and knitr [@knitr]. Detailed description of each
dataset can be found in the subsections below.

## Central Intake Call Wrap-Up Codes

On the OpenDataToronto portal, there are several datasets that reflect
the City’s effort to shelter the local homeless population. The Central
Intake Call Wrap-Up Codes Dataset [@Central_Intake_Call], stored in the
Central Intake Calls Catalogue [@Central_Intake_Call], is one of the
freshest and most detailed. Data is stored and published by the Shelter,
Support & Housing Administration since November of 2020 and refreshes on
a monthly basis. The latest refresh occurred on January 15th, 2024.

The data set provides a daily summary of the number of calls received,
the number of calls classified into distinct wrap-up codes by the nature
of its issue, and a count of calls under each wrap-up code. One of the
example wrap-up codes in the original data set was: *Code 1A - Referral
to a Sleeping/Resting Space*. The original data set includes 13 distinct
wrap-up codes; only two codes, *Code 1A — Referral to a Sleeping/Resting
Space* and *Code 2C — Information - Homelessness & Prevention Services*,
were chosen for our analysis. Code 1A and Code 2C are best suited as
measurements for positive impact the Central Intake Line could provide
because they provide a count of the number of callers provided with
directions or advice. Other non-suitable wrap-up codes include: *Code
1D - Declined Shelter/Resting Space*, *Code 4B - Disconnected - No
Outcome*, etc.

The final dataset only includes monthly cumulative data before July of
2023. The rationalization behind the action will be explained in
@sec-combined-data.

```{r}
#| label: tbl-monthly-coded
#| tbl-cap: Sample of monthly summation for Helpline Coded Calls data
#| echo: false
#| message: false

cleaned_data <- read_csv(here::here("data/analysis_data/cleaned_data.csv"))

sample_helpline <- cleaned_data |> select(Month_date, Total_Coded, Referred, Informed)

sample_helpline <- head(sample_helpline)

knitr::kable(sample_helpline, col.names = c("Month", "Calls Coded", "Referral to Shelter", "Homelessness Prevention Info"), align = c("l", "c", "c", "c"))
```

## Deaths of People Experiencing Homelessness

The Deaths of People Experiencing Homelessness Dataset
[@Homeless_Death_Count] contains monthly cumulative records of homeless
deaths. The dataset is published by Toronto Public Health [@TPH]. The
earliest data record started in January of 2017, and the latest record
ends in June of 2023.

The original dataset contains three columns: *Year of death*, *Month of
death*, and *Count*. After careful inspection of raw data, any rows that
containing an “unknown” value are excluded. Although this action created
a source of error in future analyses, this is still a necessary step
because we are plotting death counts against a timeline.

Final dataset only includes data from November 2020 to June 2023
(@tbl-death-count). The rationalization behind the action will be
explained in @sec-combined-data.

```{r}
#| label: tbl-death-count
#| tbl-cap: Sample of Cleaned Deaths of People Experiencing Homelessness Dataset
#| echo: false

sample_death_counts <- cleaned_data |> select(Month_date, Death_Count)

sample_death_counts <- head(sample_death_counts)

knitr::kable(sample_death_counts, col.names = c("Month", "Death Count"), align = c("c", "c"))
```

## Final Combined Data {#sec-combined-data}

During the data cleaning process, code was written to ensure that both
the Central Intake Call Wrap-Up Codes Dataset [@Central_Intake_Call] and
the Deaths of People Experiencing Homelessness Dataset
[@Homeless_Death_Count] are filtered to include only data between 1st
November, 2020 to 30st June, 2023.

The decision is reached by taking the common time period between the two
datasets. In the latest refresh, the Central Intake Call Wrap-Up Codes
Dataset [@Central_Intake_Call] begins on 3rd November, 2020 and end on
31st December, 2023; whereas the Deaths of People Experiencing
Homelessness Dataset [@Homeless_Death_Count] begins on January of 2017
and ends on June 2023.

The ultimate purpose of taking the common time period is to ensure that
we can have the same amount of observations for analysis. Thus a
combined dataset that kept the variables of interest in both datasets is
created (@tbl-combined).

```{r}
#| label: tbl-combined
#| tbl-cap: Sample of monthly homeless death count and Helpline Coding efforts
#| echo: false

sample_cleaned_data <- head(cleaned_data)

knitr::kable(sample_cleaned_data, col.names = c("Month", "Total Coded", "Referred to Shelter", "Homeless Info", "Death Count"), align = c("l", "c", "c", "c", "c"))

```

# Model {#sec-model}

While we can still make claims on the correlations between helpline
efforts and homeless death count without modelling, regression analysis
is still necessary to provide rigorous backbones for a claim's
statistical significance.

Here we briefly describe the Bayesian analysis model used to investigate
the correlation between homeless death counts in Toronto and the efforts
of the Helpline staff. Background details and diagnostics are included
in [Appendix -@sec-model-details].

## Model set-up

Magnitude of effect created by the helpline efforts can be estimated
through linear models. Gaussian regression model is best suited for our
analytical purposes. Other generalized linear models such as Logistic
and Multilevel are not considered due to the nature of the variables in
interest. Logistic regression model requires data that yields either an
**yes** or **no** result, whereas multilevel model require more
complexity between variables. Due to the nature of the Poisson function,
the Poisson model was considered and discarded because model results in
extremely small values with no evidently different interpretations.

### Model 1: Total Coded Model

The Total Coded model estimates the the total helpline efforts against
Toronto's homeless death count.

$$y_i|\mu_i, \sigma \sim \mbox{Normal}(\mu_i, \sigma)$$

$$\mu_i = \alpha + \phi_i$$

$$\alpha \sim \mbox{Normal}(0, 2.5)$$

$$\phi \sim \mbox{Normal}(0, 2.5)$$

$$\sigma \sim \mbox{Exponential}(1)$$

Where:

-   $y_i$ is the number of homeless death count in Toronto recorded by
    @TPH per month

-   $\phi_i$ is the total number of calls received and coded by the
    helpline per month

### Model 2: Referred Informed Model

In case of oversimplification, a second model is proposed to isolate the
effects of calls that refer individuals to shelter and calls that
provide information on homelessness prevention.

$$y_i|\mu_i, \sigma \sim \mbox{Normal}(\mu_i, \sigma)$$

$$\mu_i = \alpha + \beta_i + \gamma_i$$

$$\alpha \sim \mbox{Normal}(0, 2.5)$$

$$\beta \sim \mbox{Normal}(0, 2.5)$$

$$\gamma \sim \mbox{Normal}(0, 2.5)$$

$$\sigma \sim \mbox{Exponential}(1)$$

Where:

-   $y_i$ is the number of homeless death count in Toronto recorded by
    @TPH per month

-   $\beta_i$ is the number of homeless individuals referred to shelters
    through the helpline, per month

-   $\gamma_i$ is the number of individuals provided with information on
    homelessness prevention through the helpline, per month

We run the model in R [@citeR] using the `rstanarm` package of
@rstanarm. We use the default priors from `rstanarm`.

### Model justifications

We expect a net negative relationship between the net efforts made by
the helpline staff (Recorded by the Total_Coded) and homeless death
count, which implies that net effect of helpline calls **reduces**
homeless death count in Toronto.

Additionally, we expect a negative relationship between homeless death
and the two individual efforts: *number of homeless individuals referred
to shelter*, and *number of individuals provided with homelessness
prevention information*. Since both factors are coded efforts made by
the helpline, both factors should also help to reduce homeless death
count in Toronto.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

total_coded_model <-
  readRDS(file = here::here("models/total_coded_model.rds"))

referred_informed_model <-
  readRDS(file = 
            here::here("models/referred_informed_model.rds"))
```

Under the Gaussian regression model, @fig-total-coded illustrates a
fitted line that estimates the linear relationship between the number of
total calls coded and homeless death count in Toronto. The graph suggest
that there exist a net negative correlation between the two variable.
According to the graph, homeless death count decrease as total number of
calls coded increase. The significance of this relationship is discussed
below in in the model summary table.

```{r}
#| label: fig-total-coded
#| fig-cap: Visualization to show correlation between number of Calls Coded and Homeless Death Counts
#| echo: false
#| message: false

cleaned_data |> ggplot(aes(x = Total_Coded, y = Death_Count)) + geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = TRUE, color = "blue", linetype = "dashed", formula = "y ~ x") +
  labs(x = "Monthly Number of Calls Coded (per month)", y = "Monthly Number of Homeless Deaths")
```

The model summary table, @tbl-modelresults, displays the coefficient
values of the predictor variables against response variable: homeless
death count in Toronto. Coefficients of predictors indicate their
influence against the response variable. For example, the column on the
left shows that, under Model 1, every 1 additional helpline call coded
and decrease the homeless death count by 0.0003. The ***+p \<...***
section at the bottom of the table represents statistically significance
of the coefficient values. As indicated by the ***+p\<...*** legend, the
coefficients with 1 \*, coefficient is significant at 95% confidence
level; for \*\*, coefficient is significant at 99% confidence level, and
so on. However, as we can see from the table below, no \* exists for the
coefficients in both models.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory model of homeless death count in relation to different helpline efforts"

modelsummary(
  list(
    "Total Coded Model" = total_coded_model, "Referred and Informed Model" = referred_informed_model
  ),
  statistic = "mad",
  stars = TRUE,
  fmt = 4,
  metrics = c("RMSE", "R2")
)
```

# Discussion {#sec-discussion}

## Interpretation of Efforts

-   On average, for every increase in calls coded by the helpline, there
    is a 0.0003 unit decrease in homeless death in Toronto. There is an
    overall reduction, albeit very small, in homeless death as the
    number of total coded calls increase. This implies that the helpline
    need to take at minimum 3333 calls per month to save 1 homeless life
    in Toronto. On average, the helpline takes 14828.59 calls per month,
    meaning that 4.45 lives are saved on average.
-   4.45 is not a significant number in both a practical and statistical
    sense. To improve the return on total number of calls coded, we
    attempted to identify promising call codes that contribute to the
    net positive reduction in homeless death. Out of the 13 different
    call codes, *Referral to Shelter* and *Information on Homeless
    Prevention* are best suited as individual efforts of measurement for
    positive impacts because they provide a count of the number of
    callers provided with directions or advice. For every caller
    provided with information on homeless prevention information, there
    is an average decrease of 0.0048 death count in Toronto. On the
    other hand, for every homeless caller referred to shelter, there is
    a 0.0059 unit of average **increase** in homeless death in Toronto.
    This trend is very counter-intuitive and discouraging at first
    glance. In the case where the correlation is more significant, we
    would want to further analyze factor related to referrals to
    shelter. Accuracy on the current status of shelters, level of detail
    on instructed, and misjudgment of shelter eligibility are possible
    factors of examination. However, since the effect of this call code
    is statistically insignificant, it is also likely to be just a fluke
    of observation.

## Factors Contributing to Insignificance

Homeless life is not easy, data from @Homeless_Death_Count shows that a
total of 522 homeless individual died in Toronto between November 2020
to June 2023. The Central Intake Line is created with the intend to help
homeless individuals in need. It provides a 24/7 platform for
individuals to receive government help. The insignificance of its
influence suggests that the helpline has not fulfilled its intended
purpose, though it is also important to note that having insignificant
effect on reducing death count does not mean the helpline has not
contributed to the well-being of Toronto's homeless population in other
ways. The following two subsections discusses two of the possible
factors that contributes to the insignificance of the helpline's
influence on the reduction of homeless death in Toronto: inaccessibility
to phone lines and substitute calling options.

### Inaccessibility of Phone Lines

Difficulty in accessing the helpline may contribute to the
insignificance of the influence the helpline has on the reduction of
homeless death in Toronto. In order to access the resources provided by
the helpline, a homeless individual is required to have access to one of
the following three tools: mobile phone with valid phone plan, free
city-run phone, or payphone. Financially, the most expensive option is
the mobile phone. In 2024, the cheapest calling plan in Toronto starts
at \$15 before tax, discouraging the less wealthy proportion of the
homeless population. The cheapest alternative is the city-run free phone
lines accessible in several City buildings and Universities across
Toronto. According to @localwiki, there are currently 9 locations in
downtown Toronto offering access to free phone lines. However, the free
phone line access is only available during the operation hours of the
respective buildings, meaning that free calls are not available 24/7.
This leaves the last option, the affordable and readily available local
payphones. Currently, Toronto payphones cost 50 cents per call, which is
definitely more financially reasonable for the homeless population.
However, numerous articles through the years have pointed out the steady
decline of functional payphones in Toronto. The popularization of
smartphones in the last decade contributes to the decrease in the demand
of payphones, subsequently causing more and more abandoned or
un-repaired payphone stations. This difficulty in accessing the helpline
will decrease the number of callers, subsequently lowering the influence
of the helpline on reducing homeless death in Toronto.

### Substitute Calling Options

The Central Intake Line is one of the biggest city-run homeless
helplines in Toronto; however, it is not the only one. Using keywords
“homeless helpline Toronto”, search results such as City of Toronto Line
(311) and Assaulted Women’s Helpline also appear in the Google search
engine. The existence of multiple helplines will cause specific problems
to go to specific channels, reducing the total percent of homeless
population that choose to use the Central Intake Line inquired in this
paper. Similar to the previous factor, reduction in the total number of
calls per month reduces the average influence of the helpline on the
reduction of homeless death. Other potential factors that may contribute
to the insignificance of influence are: changes in city policies, lack
of promotion through non-internet platforms, etc.

## Weaknesses and next steps

The biggest weakness for this paper is the small sample size of the
data, there are only 32 observations in the final data used for
analysis. This suggests that resulting outputs may not accurately
represent the situation in the real world because trends might only
exist in our observed sample. Some potential next steps are:
investigating into the proposed factors that contributes to the
insignificance of helpline influence on homeless death using real-world
data and models, as well as researching further into other potential
factors.

\newpage

\appendix

# Appendix {.unnumbered}

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-total-coded-model-ppcheckandposteriorvsprior-1 we implement a
posterior predictive check. This shows a relatively close match between
the regression outcome from our data and the simulation created by the
posterior distribution. This suggests that our model is a relatively
good fit for our data.

In @fig-total-coded-model-ppcheckandposteriorvsprior-2 we compare the
posterior with the prior. We can see that the estimation parameters
shifts minimally after taking data into account. This suggests that good
prior parameters were set.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-total-coded-model-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(total_coded_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(total_coded_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

In @fig-referred-informed-model-ppcheckandposteriorvsprior-1 we
implement a posterior predictive check. This shows a relatively close
match between the regression outcome from our data and the simulation
created by the posterior distribution. This suggests that our model is a
relatively good fit for our data.

In @fig-referred-informed-model-ppcheckandposteriorvsprior-2 we compare
the posterior with the prior. We can see that the estimation parameters
shifts minimally after taking data into account. This suggests that good
prior parameters were set.

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-referred-informed-model-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(referred_informed_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(referred_informed_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-total-coded-model-stanareyouokay-1 is a trace plot. Trace plots are
often used to diagnose abnormalities of models. In
@fig-total-coded-model-stanareyouokay-1, lines are oscillating
vertically and maintains an overall horizontal trend. This suggests that
there is nothing out of the ordinary for our model.

@fig-total-coded-model-stanareyouokay-2 is a Rhat plot. Rhat plot is a
valuable tool for assessing convergence in Bayesian regression models,
helping ensure the reliability and validity of the model's inference
results. In @fig-total-coded-model-stanareyouokay-2, all data points are
close to 1. This suggest that there is nothing out of the ordinary for
our model.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-total-coded-model-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for total_coded_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(total_coded_model, "trace")

plot(total_coded_model, "rhat")
```

@fig-referred-informed-model-stanareyouokay-1 is a trace plot. Trace
plots are often used to diagnose abnormalities of models. In
@fig-referred-informed-model-stanareyouokay-1, lines are oscillating
vertically and maintains an overall horizontal trend. This suggests that
there is nothing out of the ordinary for our model.

@fig-referred-informed-model-stanareyouokay-2 is a Rhat plot. Rhat plot
is a valuable tool for assessing convergence in Bayesian regression
models, helping ensure the reliability and validity of the model's
inference results. In @fig-referred-informed-model-stanareyouokay-2, all
data points are close to 1. This suggest that there is nothing out of
the ordinary for our model.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-referred-informed-model-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm for referred_informed_model"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(referred_informed_model, "trace")

plot(referred_informed_model, "rhat")
```

\newpage

# References